---
title: "project2"
author: "Juan Piao"
date: "4/8/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## a. Use your data to plot the frontier in the mean-variance space.
```{r}
#Read your csv file:
a <- read.csv("stockData.csv", sep=",", header=TRUE)
#Convert adjusted close prices into returns:
rr <- (a[-1,4:ncol(a)]-a[-nrow(a),4:ncol(a)])/a[-nrow(a),4:ncol(a)]

#Compute the means:
means <- colMeans(rr)

#Find the covariance matrix:
cov.matrix <- cov(rr)

#Compute the vector of standard deviations:
stdev <- diag(cov.matrix)^.5

ones <- rep(1, 30)

#Compute A:
A <- t(ones) %*% solve(cov.matrix) %*% means


#Compute B:
B <- t(means) %*% solve(cov.matrix) %*% means


#Compute C:
C <- t(ones) %*% solve(cov.matrix) %*% ones


#Compute D:
D <- B*C - A^2
#Give values for E:
E <- seq(-5,5,.1)

#Compute sigma2 as a function of A,B,C,D, and E:
sigma2 <- (C*E^2 - 2*A*E +B) /D

#Plot sigma2 against E:
plot(E, sigma2, type="l", ylab=expression(sigma^2))

#Add the minimum risk portfolio:
points(A/C, 1/C, pch=19)



#Or plot E against sigma2:
plot(sigma2, E,type="l", xlab=expression(sigma^2))

#Add the minimum risk portfolio:
points(1/C, A/C, pch=19)


```

## b. Use your data to plot the frontier in the mean-standard deviation space:
### 1. Using the hyperbola method.

See the plot below. The frontier obtained using hyperbola method is in color black.

### 2. By finding two portfolios on the efficient frontier. Note: You will need to choose two values of the risk free asset.
See the plot below. The frontier obtained using by finding two portfolios on frontier is in color blue.

```{r, warning=FALSE}
#Composition:
x1 <- ( solve(cov.matrix) %*% ones ) / as.numeric( t(ones) %*% solve(cov.matrix) %*% ones )

#Mean:
m1 <- t(x1) %*% means

#Variance:
v1 <- t(x1) %*% cov.matrix %*% x1

#Portfolio 2:  (It doesn't have to be efficient, as long as it is on the frontier).
#Need to choose a value of E.  Let's say, E=0.015.
#To find x2 we use our class notes (see week 2 - lecture 1 notes):
#x2=lambda1*Sigma^-1*means + lambda2*Sigma^-1*ones
#lambda1 = (CE-A)/D and lambda2=(B-AE)/D.

E <- 0.03
lambda1 <- (C*E-A)/D
lambda2 <- (B-A*E)/D

x2=as.numeric(lambda1)*solve(cov.matrix) %*% means + 
as.numeric(lambda2)* solve(cov.matrix) %*% ones

#Mean:
m2 <- t(x2) %*% means

#Variance:
v2 <- t(x2) %*% cov.matrix %*% x2

#We also need the covariance between portfolio 1 and portfolio 2:
cov_ab <- t(x1) %*% cov.matrix %*% x2

#Now we have two portfolios on the frontier.  We can combine them to trace out the entire frontier:
#Let a be the proportion of investor's wealth invested in portfolio 1.
#Let b be the proportion of investor's wealth invested in portfolio 2.

a <- seq(-3,3,.1)
b <- 1-a

r_ab <- a*m1 + b*m2

var_ab <- a^2*v1 + b^2*v2 + 2*a*b*cov_ab
sd_ab <- var_ab^.5


plot(0, A/C, main = "Portfolio possibilities curve", xlab = "Risk (standard deviation)",
  ylab = "Expected Return", type = "n",
  xlim = c(-2*sqrt(1/C), 4*sqrt(1/C)), 
  ylim = c(-2*A/C, 4*A/C))
  
#Plot center of the hyperbola:
    points(0, A/C, pch = 19)

#Plot transverse and conjugate axes:
    abline(v = 0) #Also this is the y-axis.
    abline(h = A/C, col="red")

#Plot the x-axis:
    abline(h = 0)
    
#Plot the minimum risk portfolio:
    points(sqrt(1/C), A/C, pch=19)
    
#Find the asymptotes:
    V <- seq(-1, 1, 0.001)
    A1 <- A/C + V * sqrt(D/C)
    A2 <- A/C - V * sqrt(D/C)
    points(V, A1, type = "l", col="green")
    points(V, A2, type = "l", col="green")

#Efficient frontier:
    minvar <- 1/C
    minE <- A/C
    sdeff <- seq((minvar)^0.5, 1, by = 0.0001)
    options(warn = -1)
    y1 <- (A + sqrt(D*(C*sdeff^2 - 1)))*(1/C) 
    y2 <- (A - sqrt(D*(C*sdeff^2 - 1)))*(1/C) 
    options(warn = 0)

    points(sdeff, y1, type = "l")
    points(sdeff, y2, type = "l")


points(sd_ab, r_ab, col="blue")
```


## c. Choose a value of Rf , draw the tangent line to the efficient frontier, find the composition of the point of tangency, and the mean and variance of the point of tangency.

```{r, warning=FALSE}

#Plot the minimum risk portfolio:
    plot(sqrt(1/C), A/C,  xlim=c(0,.2), ylim=c(-.005,.05),pch=19, xlab="Portfolio standard deviation", ylab="Portfolio expected return" )
    
    
#Efficient frontier:
    minvar <- 1/C
    minE <- A/C
    sdeff <- seq((minvar)^0.5, 1, by = 0.0001)
    options(warn = -1)
    y1 <- (A + sqrt(D*(C*sdeff^2 - 1)))*(1/C) 
    y2 <- (A - sqrt(D*(C*sdeff^2 - 1)))*(1/C) 
    options(warn = 0)

    points(sdeff, y1, type = "l")
    points(sdeff, y2, type = "l")
    
#Choose risk-free return:
Rf <- 0.005

#Range of expected return:
sigma <- seq(0,.5, .001)

Rp1 <- Rf + sigma*sqrt(C*Rf^2-2*Rf*A+B) 

Rp2 <- Rf - sigma*sqrt(C*Rf^2-2*Rf*A+B) 

points(sigma, Rp1, type="l")

points(sigma, Rp2, type="l")

#===================================
#===================================
#Point of tangency:
R <- means - Rf
z <- solve(cov.matrix) %*% R

#the composition of the point of tangency
xx <- z/sum(z)
# the mean of the point of tangency
rr <- t(xx) %*% means
# the variance of the point of tangency
varr <- t(xx) %*% cov.matrix %*% xx
sdev <- varr^.5

points(sdev, rr, pch=19, col="blue")
xx
rr  # the mean of the point of tangency
sdev  # the variance of the point of tangency
```

The composition of the point of tangency is as listed above.

The mean of the point of tangency is 0.02687105.

The variance of the point of tangency is 0.02189197.

## d. Go back to the plot you constructed in project 1 and add the efficient frontier, and the tangency point from (b).

```{r}

 plot(stdev, means, col="lightblue", pch=19, cex=2, xlim=c(0, 0.3))
text(stdev, means, labels=names(means))
#points(equal_portfolio_sd, equal_portfolio_mean, col="green", pch=19, cex=2)

#minimum risk portfolio
points(minvar^0.5, minE, col="red", pch=19, cex=2)

#point of tangency
points(sdev, rr, pch=19)

#Efficient frontier
    points(sdeff, y1, type = "l")
    points(sdeff, y2, type = "l")
    
```
